---
title: "p8105_hw2_hs3239"
author: "hanfu shi"
date: "2021/10/1"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(haven)

```

problem 1
```{r}
sheet = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", na = c("",  "."),sheet = "Mr. Trash Wheel")
sheet = select(sheet, -...15, -...16, -...17)
sheet = janitor::clean_names(sheet)

sheet_omit_dumpster = filter(sheet, dumpster >0)
sheet_omit_dumpster = mutate(
    sheet_omit_dumpster,
    sports_balls = round(sports_balls)
    
  )
sheet_omit_dumpster
```

select 2018 and 2019

```{r}
rain2019 = read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14"
    ) %>%
  mutate(year = "2019") %>% 
  janitor::clean_names()
rain2019 = filter(rain2019, total >0)

rain2018 = read_excel(
    "data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    range = "A2:B14"
    ) %>%
  mutate(year = "2018") %>% 
  janitor::clean_names()
  
rain_combination = 
  bind_rows(rain2019, rain2018) %>%
  mutate(month = month.name[month]) %>% 
  select(year, month, total_inch = total)

rain_combination

```
Summary: There are 344 observations of Mr. Trash Wheel dataset after clean and 
filtration. There are 18 observations of rainfall in 2018 and 2019  after filtering missing values.The total precipitation in 2018 is               `r sum(rain2018$total) ` and the median of the sports balls in 2017 is 8




```{r}


```




Problem2

##clean pols and seperate date to three variables.
```{r}
 pols = read_csv("./data/pols-month.csv")%>%
 separate(mon, into =c("year", "month","day"), sep = c(4,7))
 pols$year =as.integer(pols$year)
 pols = separate(pols, month, into = c("useless","month"), sep = 1)
 pols$month =as.integer(pols$month)
 pols$month =month.name[pols$month]
   pols=select(pols, -useless)
 

```

##create president
```{r}
pols = mutate(
  pols, 
  president = pols$prez_gop + pols$gov_gop+ pols$sen_gop+ pols$rep_gop+ pols$prez_dem+ pols$gov_dem+pols$sen_dem+pols$rep_dem
 )

pols = select(pols, -prez_dem, -prez_gop, -day)

 head(pols)
```

##snp
```{r}
 snp = read.csv("./data/snp.csv")%>%
 separate(date, into =c("month", "day","year"), sep = c(1,-4))

 snp$year =as.integer(snp$year)
 
 snp$month =as.integer(snp$month)
 snp$month =month.name[snp$month]
 snp = select(snp, -day)%>%
   relocate(year,month)%>%
   arrange(year,month)
 
head(snp)
```


```{r}
unemp = 
  read.csv(file = "./data/unemployment.csv") %>%
  
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "percentage"
  ) %>%
janitor::clean_names()%>%
  mutate(month = month.name[match(pull(., month), month.abb)],
         year = as.integer(year)) 
 

head(unemp)

```


combine three tables

```{r}
combine_three = left_join(pols, snp, by = c("year", "month")) %>% 
  left_join(unemp, by = c("year", "month")) %>% 
  drop_na(close, percentage)

head(combine_three)
```
Summary: pols dataset has 822 observations and 3 variables. snp dataset has 787 observatons and 3 variables. unempoly dataset has 816 data set. There are
786 observations and 12 variables after combination. 


P3

```{r}
PBaby = 
  read.csv(file = "./data/Popular_Baby_Names.csv")%>%
  janitor::clean_names()%>%
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    child_s_first_name = str_to_lower(child_s_first_name),
    ethnicity = replace(ethnicity, ethnicity == "asian and paci", "asian and pacific islander"),
    ethnicity = replace(ethnicity, ethnicity == "black non hisp", "black non hispanic"),
    ethnicity = replace(ethnicity, ethnicity == "white non hisp", "white non hispanic")
    ) %>% distinct()

head(PBaby)
```

Olivia rank table
```{r}
Olivia = filter(PBaby, gender == "female", child_s_first_name == "olivia")%>%
  select(year_of_birth, ethnicity, rank)%>%
  arrange(rank)%>% 
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "rank")

knitr::kable(Olivia, format = "html", caption = "Olivia rank through years")


```


```{r}
male_name =filter(PBaby, gender == "male", rank == "1")%>%
  select(year_of_birth, ethnicity, child_s_first_name)%>%
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "child_s_first_name")

knitr::kable(male_name, format = "html", caption = "Olivia rank through years")


```


